---
title: "Homework7"
author: "Rohan Sadale"
date: "March 10, 2016"
output: pdf_document
---

```{r}
library(alr4)
```

### 6.1

```{r}
m1 <- lm(lifeExpF ~ 1, UN11)
m2 <- lm(lifeExpF ~ group, UN11)
anova(m1, m2)
```

The summary shows that p-value is almost 0. Thus group should not be removed.

### 6.2

For F-tests to be done, NH should be special case of AH (such as setting one of the $\beta$ parameter to zero). 

### 6.3

```{r}
m1 <- lm(lifeExpF ~ group, UN11)
m2 <- lm(lifeExpF ~ group + log(ppgdp), UN11)
anova(m1, m2)
```

The summary says that as p-value is almost 0, we should not remove log(ppgdp).

### 6.4

#### 6.4.1 

Common intercept and different slopes in each group. The absence of group removes all the different intercepts.

#### 6.4.2

```{r}
m1 <- lm(lifeExpF ~ log(ppgdp) + group:log(ppgdp), UN11)
m2 <- lm(lifeExpF ~ group + log(ppgdp) + group:log(ppgdp), UN11)
anova(m1, m2)
```

As p-value isn't 0, we should remove the interaction term.

### 6.5

#### 6.5.1

```{r}
m1 <- lm(lifeExpF ~ group + log(ppgdp), UN11)
summary(m1)
```

From the p-value of groupother we can say that it isn't significantly different from the intercept(which is in this case oecd)


#### 6.5.2

```{r}
linearHypothesis(m1, "groupother - groupafrica")
```

From the p-value we can say that intercept of other and africa aren't different.

### 6.6

```{r}
data(fuel2001)
fuel2001$Dlic = 1000*fuel2001$Drivers/fuel2001$Pop
fuel2001$Fuel = 1000*fuel2001$FuelC/fuel2001$Pop
fuel2001$Income = fuel2001$Income/1000

m1 <- lm(Fuel ~ 1, fuel2001)
m2 <- lm(Fuel ~ Tax+Dlic+Income+log(Miles), fuel2001)
anova(m2, m1)
```

From the summary we can say that model with all the regressors has p-value of almost 0 (significant). Thus we don't need to do any futher testing such as removing regressors.

### 6.7

#### 6.7.1

```{r}
data(fuel2001)
fuel2001$Dlic = 1000*fuel2001$Drivers/fuel2001$Pop
fuel2001$Fuel = 1000*fuel2001$FuelC/fuel2001$Pop
fuel2001$Income = fuel2001$Income/1000

m1 <- lm(Fuel ~ Tax+Dlic+Income+log(Miles), fuel2001)
m2 <- lm(Fuel ~ log(Miles)+Tax+Dlic+Income, fuel2001)

anova(m1)
anova(m2)
```

As anova() provides tests in sequential order, there is difference between the anova for m1 and m2.


#### 6.7.2

```{r}
data(fuel2001)
fuel2001$Dlic = 1000*fuel2001$Drivers/fuel2001$Pop
fuel2001$Fuel = 1000*fuel2001$FuelC/fuel2001$Pop
fuel2001$Income = fuel2001$Income/1000

m1 <- lm(Fuel ~ Tax+Dlic+Income+log(Miles), fuel2001)
m2 <- lm(Fuel ~ log(Miles)+Tax+Dlic+Income, fuel2001)

Anova(m1)
Anova(m2)
```

The regressors which are added at the last in Type 1 anova have same estimates as respective those in Type 2 anova. 