---
title: "Homework1"
author: "Rohan Sadale"
output: html_document
date: "Jan 28, 2016"

---
```{r, message=FALSE}
library(alr4)

```


#### 1.1.

**1.1.1**

* Predictor: ppgdp
* Response: fertility


**1.1.2**

```{r}
with(UN11, plot(fertility ~ ppgdp))
```


The plot is right skewed. Hence Simple Linear Regression cannot be a good fit for this graph. The straight line mean function is not possible for this graph.


**1.1.3**

```{r}
with(UN11, plot(log(fertility) ~ log(ppgdp)))
model <- lm(log(fertility) ~ log(ppgdp), data = UN11)
abline(model)
```


Yes, as we can see from the plot, Simple Linear Regression seem plausible for a summary of this graph.


#### 1.2

```{r}
meanData <- with(wblake, tapply(Length, Age, mean))
print(meanData)

varianceData <-with(wblake, tapply(Length, Age, var))
print(varianceData)

plot(1:8, meanData)
```


The graph shows a rising trend in average length of fish as it grows older, and a simple linear regression seems plausible for this graph. The graph has same pattern as compared to fig 1.5 


```{r}
plot(1:8, sqrt(varianceData))
```


Though a straight line parallel to Age axis is possible on the plot, but the variance and standard deviation of the fishes at age 3 and age 8 will deviate significantly from a straight line. Thus, the plot is not a null plot.



#### 1.3

**1.3.1**
There appears to be no dependence of average soil temperature on the number of months. The plot appears to be null plot as it doesn't show any particular important characteristics.


**1.3.2**
```{r fig.width=12, fig.height=3}
with(Mitchell, plot(Temp~Month))
```


As the width changes, we are able to see a temporal pattern on data. We can say that the soil temperature falls in Winter and rises in summer.


#### 1.4

```{r}
with(oldfaith, plot(Interval~Duration))
model2 <- lm(Interval~Duration, data = oldfaith)
abline(model2)
plot(residuals(model2))
abline(0,0)
```

The mean is almost linear, variance is almost constant and the residual plot can be a null plot. Thus, we can summarize the graph by simple linear regression.


#### 1.5

```{r}
plot(water)
```

* From the plot, we can say that Year is highly uncorelated to all other variables. 
* We can also notice that the variables OPBPC, OPRC, OPSLAKE and BSAAM are corelated with each other. Similarly, the variables - APMAM, APSAB and APSLAKE appear to be corelated with each other.
* BSAAM is more corelated with OPBPC, OPRC and OPSLAK, while less corelated with APMAM, APSAB and APSLAKE.
* We can also see that there is atleast one seperated point (outlier or leverage point) in each plot.

#### 1.6

* We can see that quality and clarity, quality and helpfulness are highly corelated to each other. Clarity and helpfulness are a bit less corelated to each other as compared to former two.
* Easiness is very loosly corelated to helpfulness, clarity and quality. 
* There seems no corelation between raterInterest and all other variables.